---
title: "explore_fission_fusion_features"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dichromat)

# name <- readline(prompt = "Enter name:")
# 
# if(grepl('a|A', name)){
#   basedir <- '~/Dropbox/hyenas/hyena_fission_fusion/' # <-- CHANGE TO THE MAIN PROJECT FOLDER
#   codedir <- '~/Dropbox/code_ari/hyena_fission_fusion/' # <-- CHANGE TO THE DIRECTORY WHERE CODE REPOSITORY IS
# }else if(grepl('e|E', name)){
  basedir <- 'C:/Users/strau/Dropbox/hyena_fission_fusion/' # <-- CHANGE TO THE MAIN PROJECT FOLDER
  codedir <- 'C:/Users/strau/Dropbox/Documents/Research/Partial_projects/hyena_fission_fusion/' # <-- CHANGE TO THE DIRECTORY WHERE CODE REPOSITORY IS
# }else{
#   cat('Name not recognized')
# }

#SUBDIRECTORIES
datadir <- paste(basedir, 'data/raw', sep = '')
processeddir <- paste(basedir, 'data/processed', sep = '') #this is where the output will be saved (the fission fusion events RData file)

#FUNCTIONS
source(paste0(codedir,'/hyena_functions.R'))

#PROCESSED DATA
load(file = paste0(processeddir,'/fission_fusion_features.RData'))
load(file = paste0(processeddir,'/fission_fusion_events.RData'))
load(file = paste0(datadir,'/hyena_xy_level1.RData'))


```

## Fission-fusion events

Fission-fusion events are identified as cases where two individuals get within 100m of each other. 
Once an event is identified, data are collected from the following periods:  
* Before - When individuals are between 300 and 200m apart  
* During - When individuals are within 200m of each other  
* After - When individuals are between 200m and 300m of each other  

Here is some basic exploration of this data

```{r general exploration, echo = FALSE}

good.idxs <- which(events$start.exact & events$end.exact)
examples <- good.idxs[45:49]

hist((events$duration.together + events$duration.fusion + events$duration.fission)/60, breaks = 500, main = 'Duration of fission-fusion events', xlab = 'Minutes together')

par(mfrow = c(1,1))
plot_events(indices = examples, together.seqs = events, xs, ys)
```

Our goal is to characterize fission-fusion events and categorize them into a set
of biologically meaningful categories. To do this, we decided to start by looking
at the distance between the pair of individuals over time during each event. 

Fission-fusion events tend to have a canonical structure where there is first a 
decrease in the dyadic distance (as the two hyenas get closer together; "fusion phase"), then a 
relatively stable period where the hyenas spend time close together ("together phase"), followed by
an increase in the dyadic distance as they part ways ("fission" phase). 

Based on this idea, we decided to fit all fission-fusion events to this canonical 
structure using a 3-part piecewise function (further details on this are available
but seemed distracting). Although the fit isn't always perfect, this approach does
do a reasonable job at capturing the structure of the events.

Here are some plots of example events with dyadic distance in black and the fitted 
phases in red. 

```{r plot a single event, echo = FALSE}
par(mfrow = c(1,2))
plot_events(indices = examples[1], events, xs, ys)
plot_canonical_shape(rows = examples[1], dyad.dists = dyad.dists, events)

plot_events(indices = examples[2], events, xs, ys)
plot_canonical_shape(rows = examples[2], dyad.dists, events)

plot_events(indices = examples[3], events, xs, ys)
plot_canonical_shape(rows = examples[3], dyad.dists, events)
```

Intuitively we can think of events as varying in which individuals are moving and 
which are stationary in each phase. Here we plot the displacement of one individual vs 
the other individual in each of the phases separately. 


```{r plot distance travelled for each individual, echo = FALSE}

plot(events$disp.fusion.i[good.idxs], events$disp.fusion.j[good.idxs], xlim = c(0,700), ylim = c(0, 700),
     xlab = 'Distance travelled by A', pch = 19, col = scales::alpha('black', 0.3), cex = 0.8, ylab = 'Distance travelled by B', main = 'Fusion phase')

plot(events$disp.together.i[good.idxs], events$disp.together.j[good.idxs], xlim = c(0,700), ylim = c(0, 700),
     xlab = 'Distance travelled by A', pch = 19, col = scales::alpha('black', 0.3), cex = 0.8, ylab = 'Distance travelled by B', main = 'Together phase')

plot(events$disp.fission.i[good.idxs], events$disp.fission.j[good.idxs], xlim = c(0,700), ylim = c(0, 700),
     xlab = 'Distance travelled by A', pch = 19, col = scales::alpha('black', 0.3), cex = 0.8, ylab = 'Distance travelled by B', main = 'Fission phase')





```

Especially in the fission and fusion phases, we see two clear clusters where one 
individual is stationary and the other is moving. We also see a large scattering 
of points representing cases where both individuals are moving substantially. 




```{r plot angle between trajectories of two individuals, echo = FALSE}

min.move <- 50

both.move.fusion <- which(events$disp.fusion.i >= min.move & events$disp.fusion.j >= min.move)
both.move.together <- which(events$disp.together.i >= min.move & events$disp.together.j >= min.move)
both.move.fission <- which(events$disp.fission.i >= min.move & events$disp.fission.j >= min.move)

hist(events$angle.fusion[both.move.fusion], main = 'Angles during fusion', breaks = 50, xlab = 'Angle (radians)')
hist(events$angle.together[both.move.together], main = 'Angles during together', breaks = 50, xlab = 'Angle (radians)')
hist(events$angle.fission[both.move.fission], main = 'Angles during fission', breaks = 50, xlab = 'Angle (radians)')



pal <- colorRampPalette(c('blue', 'black', 'goldenrod1'))
cols <- pal(320)

# Same plots as before but colored by angle
plot(events$disp.fusion.i[good.idxs], events$disp.fusion.j[good.idxs], xlim = c(0,700), ylim = c(0, 700),
     xlab = 'Distance travelled by A', pch = 19, cex = 0.8, ylab = 'Distance travelled by B', main = 'Fusion phase',
     col = cols[ceiling(events$angle.fusion[good.idxs] * 100)])

plot(events$disp.together.i[good.idxs], events$disp.together.j[good.idxs], xlim = c(0,700), ylim = c(0, 700),
     xlab = 'Distance travelled by A', pch = 19, cex = 0.8, ylab = 'Distance travelled by B', main = 'Together phase',
     col = cols[ceiling(events$angle.together[good.idxs] * 100)])

plot(events$disp.fission.i[good.idxs], events$disp.fission.j[good.idxs], xlim = c(0,700), ylim = c(0, 700),
     xlab = 'Distance travelled by A', pch = 19, cex = 0.8, ylab = 'Distance travelled by B', main = 'Fission phase',
     col = cols[ceiling(events$angle.fission[good.idxs] * 100)])

```






